\sys: Light-touch Data-Parallel Shell Processing

### 2020-10-05 paper/draft.tex
Add minted in the rest of the code

\begin{lstlisting}[language=sh, numbers=none, escapeinside={($}{$)}]
 cat f1 f2 | grep "foo" > f3 && sort f3 
\end{lstlisting}
% # Replace sort with sth better

@@

\begin{lstlisting}[language=sh, numbers=none, escapeinside={($}{$)}]
 mkfifo t1 t2
 grep "foo" in1 > t1 & 
 grep "foo" in2 > t2 &
 cat t1 t2
\end{lstlisting}

@@

\heading{Dangling FIFOs and Zombie Producers}
%
%% \TODO{(This problem is a nice followup for the one above) Problem:
%%   Dataflow graph termination. Illustration: Show the example of head
%%   deadlock and explain that this is inherent behaviour in the shell.
%%   (Solution: emptying pipelines or terminating processes) In order for
%%   things to terminate gracefully, dish wraps all commands in order to
%%   always consume all of their input by sending it down to
%%   /dev/null. Warning: This means that all commands in the pipeline
%%   must terminate gracefully if all of their output is consumed (if a
%%   process dies unexpectedly, noone will consume its input. We leave
%%   addressing that and other monitoring questions for future work). We
%%   evaluate the overhead of emptying all outputs in section X.}
%
% \nv{I think all of these titles should be the problem/challenge, not the solution}
% A process writing to a pipe needs to be notified when its consumer exits early before consuming all its output.
Under normal operation, a command exits after it has produced and sent all its results to its output channel.
If the channel is a pipe and its reader exits early, the command is notified to stop writing early.
In \unix, this is achieved by an out-of-band error mechanism: the
operating system delivers a \ttt{PIPE} signal to the producer,
notifying it that the pipe's consumer has exited.  This is different
from errors for other system calls
%% return an error
and unusual compared to non-\unix systems\footnote{For example,
  Windows indicates errors for \ttt{WriteFile} using its return
  code---similar to \ttt{DeleteFile} and other Win32 functions.}
primarily because pipes and pipelines are at the heart of \unix.
% write errors are communicated differently 
Unfortunately though, if a pipe has not been opened for writing yet
% ---possible for named pipes---
  \unix cannot signal this condition.
% \kk{I think that it is also possible for anonymous, e.g. \ttt{cat f1 - | head -n 1}}
% \nv{TODO! Will think about this later}
Consider the following script:
% emitted code from \sys's earlier versions:
%% \kk{Proposed alternative: ``Consider the following shell script:''}
\begin{lstlisting}[language=sh, numbers=none]
  mkfifo fifo1 fifo2
  cat file-chunk1.txt > fifo1 &
  cat file-chunk2.txt > fifo2 &
  cat fifo1 fifo2 | head -n 1 & wait
\end{lstlisting}
\noindent
In the code above, \ttt{head} exits early causing the last \ttt{cat} to exit before opening \ttt{fifo2}.
As a result, the second \ttt{cat} never receives a \ttt{PIPE} signal that its consumer exited---after all, \ttt{fifo2} never even had a consumer.
This, in turn, leaves it unable to make progress, as it is both blocked and unaware of its consumer exiting.
Coupled with \ttt{wait} at the end, the entire snippet reaches a deadlock.

>>>>

% \begin{lstlisting}[language=sh, numbers=none, escapeinside={($}{$)}]
%  cat f1 f2 | grep "foo" > f3 && sort f3 
% \end{lstlisting}
\begin{minted}[fontsize=\footnotesize]{bash}
 cat f1 f2 | grep "foo" > f3 && sort f3 
\end{minted}
% # Replace sort with sth better

@@

% \begin{lstlisting}[language=sh, numbers=none, escapeinside={($}{$)}]
%  mkfifo t1 t2
%  grep "foo" in1 > t1 & 
%  grep "foo" in2 > t2 &
%  cat t1 t2
% \end{lstlisting}
\begin{minted}[fontsize=\footnotesize]{bash}
 mkfifo t1 t2
 grep "foo" in1 > t1 & 
 grep "foo" in2 > t2 &
 cat t1 t2
\end{minted}

@@

\heading{Dangling FIFOs and Zombie Producers}
%
%% \TODO{(This problem is a nice followup for the one above) Problem:
%%   Dataflow graph termination. Illustration: Show the example of head
%%   deadlock and explain that this is inherent behaviour in the shell.
%%   (Solution: emptying pipelines or terminating processes) In order for
%%   things to terminate gracefully, dish wraps all commands in order to
%%   always consume all of their input by sending it down to
%%   /dev/null. Warning: This means that all commands in the pipeline
%%   must terminate gracefully if all of their output is consumed (if a
%%   process dies unexpectedly, noone will consume its input. We leave
%%   addressing that and other monitoring questions for future work). We
%%   evaluate the overhead of emptying all outputs in section X.}
%
% \nv{I think all of these titles should be the problem/challenge, not the solution}
% A process writing to a pipe needs to be notified when its consumer exits early before consuming all its output.
Under normal operation, a command exits after it has produced and sent all its results to its output channel.
If the channel is a pipe and its reader exits early, the command is notified to stop writing early.
In \unix, this is achieved by an out-of-band error mechanism: the
operating system delivers a \ttt{PIPE} signal to the producer,
notifying it that the pipe's consumer has exited.  This is different
from errors for other system calls
%% return an error
and unusual compared to non-\unix systems\footnote{For example,
  Windows indicates errors for \ttt{WriteFile} using its return
  code---similar to \ttt{DeleteFile} and other Win32 functions.}
primarily because pipes and pipelines are at the heart of \unix.
% write errors are communicated differently 
Unfortunately though, if a pipe has not been opened for writing yet
% ---possible for named pipes---
  \unix cannot signal this condition.
% \kk{I think that it is also possible for anonymous, e.g. \ttt{cat f1 - | head -n 1}}
% \nv{TODO! Will think about this later}
Consider the following script:
% emitted code from \sys's earlier versions:
%% \kk{Proposed alternative: ``Consider the following shell script:''}
% \begin{lstlisting}[language=sh, numbers=none]
%   mkfifo fifo1 fifo2
%   cat file-chunk1.txt > fifo1 &
%   cat file-chunk2.txt > fifo2 &
%   cat fifo1 fifo2 | head -n 1 & wait
% \end{lstlisting}
\begin{minted}[fontsize=\footnotesize]{bash}
  mkfifo fifo1 fifo2
  cat file-chunk1.txt > fifo1 &
  cat file-chunk2.txt > fifo2 &
  cat fifo1 fifo2 | head -n 1 & wait
\end{minted}
\noindent
In the code above, \ttt{head} exits early causing the last \ttt{cat} to exit before opening \ttt{fifo2}.
As a result, the second \ttt{cat} never receives a \ttt{PIPE} signal that its consumer exited---after all, \ttt{fifo2} never even had a consumer.
This, in turn, leaves it unable to make progress, as it is both blocked and unaware of its consumer exiting.
Coupled with \ttt{wait} at the end, the entire snippet reaches a deadlock.
